---
title: "Olive data"
subtitle: "Ordering displays"
fontsize: 9pt
geometry: margin=.75in
output:
  html_document:
    df_print: paged
  pdf_document: default
header-includes:
- \usepackage{graphicx}
- \usepackage{color}
graphics: yes
classoption: letter
---
```{r setup, echo=FALSE, message=FALSE, error = FALSE, warning=FALSE}
library(knitr)
library(loon)
```


This contains measurements on the fatty acid content of 572 different Italian olive oils; eight different fatty acids are measured.  
The olive oils come from one of nine different olive growing regions in Italy. 

This data are easily available from the `loon` package:


```{r, tidy=TRUE}
library(loon) 
# The first three rows of which are
head(olive, 3)
```

```{r, echo = FALSE}
head(olive) 
```
    
## Side by side boxplots for fatty acid 'oleic' content in 9 regions of Italy


```{r, echo=TRUE, message=FALSE, error = FALSE, warning=FALSE}
    library(colorspace)
    cols <- rainbow_hcl(9) # Use these colours.
```
     
```{r}
# separate the area and oleic acid variable
area <- olive$Area
oleic <- olive$oleic
df <- data.frame(area,oleic)
par(mar=c(8,4,4,10))
boxplot(oleic~area, data=df, col=cols, main="Oleic acid content found in 9 areas", las=2, xlab="")
mtext("Area", side=1)
```
    
## Comparisons based on Eulerians of Oleic acid in 9 regions

```{r, echo=FALSE, message=FALSE, error = FALSE, warning=FALSE}
  library(PairViz)

```
```{r}
  areas <- with(olive, split(oleic,Area))
  areaNames <- names(areas)
  ord <- eulerian(9)
  ord
```
 
  - 36 pairwise comparisons. 
    
```{r}
  par(mar=c(9,4,4,10))
  boxplot(areas[ord], col=cols[ord], main="Oleic acid content found in 9 regions", las=2, ylab="oleic acid")
  mtext("Area", side=1)
```
    
  
  We can see the North Apulia region appears more than once before all the other nodes appear.     
    
## Comparisons based on Hamiltonian of Oleic acid in 9 regions

We want to construct a hamiltonian, so that all groups appear once followed by another. 
```{r}
        ordH <- hpaths(9, matrix=FALSE)
        ordH
```
  As we can see that all the nodes appear in a pairwise comparison before repeating again.
              
```{r}
       par(mar=c(9,4,4,10))       
       boxplot(areas[ordH], col=cols[ordH], main="Oleic acid content found in 9 regions", las=2, ylab="oleic acid")
       mtext("Area", side=1)
```
  
  
  We see 9 different colours hence 9 different regions before another group for comparison appears. 
  

## Pairwise T test for Oleic acid data

Using the r function for producing pairwise t test values. 
```{r}
        # Get p-values for each area for oleic
        test <- with(olive, pairwise.t.test(oleic, Area))
        pvals <- test$p.value
      
        # Create a distance matrix
        groups <- length(areaNames)
        weights <- matrix(0, nrow=groups, ncol=groups)
        
        rownames(weights) <- areaNames
        colnames(weights) <- areaNames
        weights[2:groups, 1:(groups-1)] <- pvals
        
        # Create a matrix where diagonal is 0 and the upper and lower half are equal. 
        diag(weights) <- 0
        
        for (i in 1:(groups-1)){
          for (j in (i+1):groups){
            weights[i,j] <- weights[j,i]
          }
        }
        
        round(weights, 4)
        
        # Most significant difference would have lowest p-value so we need a greedy algorithm which goes from lowest to highest edge weights. 
        low2high_ord <- eulerian(weights); colnames(weights)[low2high_ord]
        low2high_ord
         
```
```{r}
        par(mar=c(7,7,2,2))       
        boxplot(areas[low2high_ord], col=cols[low2high_ord], main="Oleic acid content found in 9 regions", las=2)
```
  
  
  Yes, since the lowest p-value is selected from each node and no edge (pairwise comparison) is repeated. The ordering is successful because no pairwise comparison is repeated and using a Greedy eulerian from lowest to highest choosing the edge from the node with the smallest p-value and so on without repeating the edges from the node and so no comparison is repeated. 


```{r}
      low2high_ord[1:8]
      par(mar=c(7,7,2,2))  
      boxplot(areas[low2high_ord], col=cols[low2high_ord], main="Oleic acid content found in 9 regions", las=2)
```


## Comparison of each growing area in the south and each area in Sardinia

The 4 Southern areas are Sicily, South-Apulia, North-Apulia and Calabria. The 2 Sardinia areas are Inland-Sardinia and Coast-Sardinia. 
      
```{r}
      south_nodes <- c("Sicily", "South-Apulia", "North-Apulia", "Calabria")
      sardinian_nodes <- c("Inland-Sardinia", "Coast-Sardinia")
      g <- bipartite_graph(south_nodes, sardinian_nodes)
      plot(g, "circo")
```
            
Construct Eulerian using the graph from part above.
```{r}
        gEulOrd <- eulerian(g)
        gEulOrd
```
  
   Construct boxplot using this eulerian order
```{r}
        par(mar=c(9,4,4,10)) 
        names(cols) <- areaNames
        boxplot(areas[gEulOrd], col=cols[gEulOrd], main="Oleic acid comparison between South and Sardinia region", las=2)
```